Topological Equivariant Artist Model for Visualization Library Architecture

### 2021-10-04 paper/main.tex
F, K, M, A,

@@

\section{Artist} 
The \textcolor{artist}{Artist} $\mathcal{\vartist}$\footnote{We call this transformation the artist because the Artist object in Matplotlib\cite{hunterMatplotlib2DGraphics2007} is responsible for converting data to a prerender graphical object} is a transformation from data  $\mathcal{\dtotal}$ to graphic $\mathcal{\gtotal}$.

@@

We formally define data and graphic as the categories $\mathcal{\dtotal}$ and $\mathcal{\gtotal}$. To explain these categories, we first define the category of open sets $\mathcal{\dbase}$ and the category of fiber spaces $\mathcal{\dfiber}$.

\begin{definition} The category $\mathcal{\dbase}$ consists of 
  \begin{itemize}
    \item \textit{objects} all open sets $U_{i} \in \{U\}$ in \dbase, including the empty set $\varnothing$ and the maximum set \dbase
    \item \textit{morphisms} inclusion maps from a subspace to a larger space $\iota: U_{i} \hookrightarrow U_{j}$, which is equivalent to $U_{i} \subseteq U_{j}$ and  
  \end{itemize} 
\end{definition}

@@

\begin{definition} For each open set object $U_{i}$ of category $\mathcal{\dbase}$, there is a set $\Gamma(U_i, \dtotal)$. The elements of this set are the sections $\{\tau^{j}:U_i \rightarrow \dfiber_{U_{i}}\}_{j\in I}$ of \dtotal\ over open set $\U_{i}$. The sections $\Gamma(\dbase, \dtotal)$ are called the global sections of $\dtotal$. \cite{SheafMathematics2021,spanier1989algebraic}.
\end{definition}

Each section is how we represent a unique dataset, where the set of sections $\Gamma$ is the set of all datasets that have the same data fields in the \dfiber\ and continuity as encoded in\dbase\. For example, database tables that have the same schema and discrete index, as illustrated in Spivak \cite{spivakSIMPLICIALDATABASES}.

\begin{definition} The fiber space category $\mathcal{\dfiber}$ is a category of sets \textbf{Set}. This means its objects are sets, and its morphisms are total functions\cite{barrCategoryTheoryComputing}
\end{definition}

\begin{definition} The monad $T$ is [a,b,c]

@@

For example, these functions could be the binary operations, group actions, and measurement scales discussed in \autoref{sec:related-work:equivariance}. These functions could also be monoid actions, which are a way of applying partial order relations to data\cite{fongInvitationAppliedCategory2019}, such as to build multi-ranked indicators\cite{bruggemannRankingPrioritizationMultiindicator2011}
}

\begin{definition} The category $\mathcal{\dtotal}$ consists of 
  \begin{itemize}
    \item\textit{object} the space of continuous\footnote{or piecewise continuous}sections $\sheaf: U_{i} \rightarrow \Gamma(U_{i}, \dtotal) $ which is a sheaf\cite{spanier1989algebraic,SheafMathematics2021}. 
    \item\textit{morphisms} bundle maps E, E, fiber to itself and is the fiber map allowable, functions F to itself forms a monoid (one object category), F and it's allowable maps form a category, monoid could be monotonic functions, group actions, extending that to bundle E, interested in E to itself, which when restricted to fiber are category maps of fiber to itself, maps from sheaf to itself, that when restricted to fiber gives you fiber to itself, given bundle map E->E, then you can show that you have a sheaf map... Given E->E, and F->F, then there's a map from one sheaf to another. bundle map induces on level of sections induces on sheafs -> part of the definition of the presheaf

  \end{itemize}
\end{definition
The functions on the data

\note{equivariance is a special case of functions}
\note{actions are a special case of actions}
\begin{definition} category of bundle maps, which include category of fiber maps, what are allowable of fiber to itself, ex translation and scales only ..., objects are a fiber, maps are fiber to fiber defined as their own category, E to V fibers changes

zooming is inclusion maps on both sides 
\end{definition}

The objects of the data category $\mathcal{\dtotal}$ and graphic category $\mathcal{\gtotal}$ are the sheaves $\Gamma(\dbase, \dtotal)$ and $\Gamma(\gbase, \gtotal)$. An alternate notation is $\sheaf(\dtotal)$ and $\sheaf(\gtotal)$. The artist is a morphism of sheafs $\sheaf(\dtotal)$ to $\sheaf(\gtotal)$, which means it is by definition a natural transform\cite{SheafMathematics2021}. As mentioned in \autoref{sec:related-work:equivariance:category}

that the artist $\vartist$ transforms functions that act on data $f:\dtotal\rightarrow \dtotal$ to functions that act on graphics $\g: \gtotal\rightarrow \dtotal$ in a way in which the following diagram commutes 
\begin{tikzcd}
  \mathcal{\dtotal} \arrow[r, "\vartist"] \arrow[d, "f"] & \mathcal{\gtotal} \arrow[d, "g"] \arrow[d, "\ca"] \\
  \mathcal{\dtotal} \arrow[r, "\vartist"]                & \mathcal{\gtotal}                                
\end{tikzcd}
meaning that, as described in \autoref{sec:related-work:equivariance}, the artist \vartist is constructed such that $\vartist(f(\dtotal)) = g(\vartist(\dtotal))$, meaning a transformation of the data has a consistent equivalent transformation of the graphic. In \autoref{sec:artist:construction}, we show a method of constructing the artist such that continuity and equivariance are preserved. We then use this formulation to guide the development of visualization library components in \autoref{sec:case-study}.

@@

\subsection{Graphic to Data: \vindex}
\begin{equation}
  \begin{tikzcd}
      \dtotal \arrow[d, "\pi"'] & \gtotal \arrow[d, "\pi"'] \\
      \dbase                   & \gbase \arrow[l, "\vindex"']
  \end{tikzcd}
  \label{eq:math:graphic:vindex}
\end{equation}

@@

\subsection{Visual Bundle \vtotal}
\begin{equation}
  \begin{tikzcd}[ampersand replacement=\&]
      \vfiber \arrow[r, hook] \& \vtotal \arrow[d, "\pi"'] \\
                        \& \dbase \arrow[u, "\vsection"', bend right]
  \end{tikzcd}
\end{equation}

\subsection{Data to Visual Encodings: \vchannel} %look at what K&S call these stages
Visual encoding functions \vindex\ are functors because they are expected to preserve structure on the data and visual side, as discussed in \autoref{sec:related-work:equivariance}.

>>>>

Throughout this paper, we denote fiber bundles with the tuple $(\textcolor{total}{total space}, \textcolor{base}{base space}, \pi, \textcolor{fiber}{fiber space})$

@@

\section{Artist} 
The \textcolor{artist}{Artist} $\mathcal{\vartist}$\footnote{We call this transformation the artist because the Artist object in Matplotlib\cite{hunterMatplotlib2DGraphics2007} is responsible for converting data to a prerender graphical object} is a functor from data $\mathcal{\dtotal}$ to graphic $\mathcal{\gtotal}$.}

@@

We formally define data and graphic as the categories $\mathcal{\dtotal}$ and $\mathcal{\gtotal}$. We develop these categories as analogues to fiber bundle $(\dtotal, \dbase, \pi, \dfiber)$, introduced in \autoref{sec:related-work:fiber-bundles}, and fiber bundle $(\gtotal, \gbase, \pi, \gfiber)$. Throughout this section, we assume that $\mathcal{E}$ is defined in terms of a fixed continuity and fiber space, for example a class of images or tables with the same schema. Similarly we assume that $\mathcal{H}$ is a stand in for a specific type of graphic, such as a scatter or line plot.

As introduced in \autoref{sec:related-work:fiber-bundles}, sets of open sets form the basis of topological spaces. We propose that we can express how the elements in our dataset are glued together by modeling that continuity as a topological space \dbase, and we express the constraints on \dbase by defining the analogues category $\mathcal{\dbase}$.

\begin{definition} The category of open sets $\mathcal{\dbase}$ consists of
  \label{def:category:K} 
  \begin{itemize}
    \item \textit{objects} all open sets $U_{i} \in \{U\}$ in the topological space\cite{bradleyTopologyCategoricalApproach2020, } $(\dbase, \mathscr{T})$), including the empty set $\varnothing$ and the maximum set \dbase. 
    \item \textit{morphisms} inclusion maps from a subspace to a larger space $\iota: U_{i} \hookrightarrow U_{j}$, which is equivalent to $U_{i} \subseteq U_{j}$ and  
  \end{itemize} 
\end{definition}

@@

\begin{definition} 
  \label{def:category:F}
  The category $\mathcal{\dfiber}$ is a subcategory of an arbitrary category $\mathcal{C}$. $\mathcal{\dfiber}$ is a monoidal category, meaning it has one object \texttt{\dfiber} and a functor\footnote{The functor $\otimes$ is a bifunctor because it takes as input the product of two categories \cite{nlab:bifunctor}} $\otimes:\mathcal{\dfiber}\times{\mathcal{\dfiber}}\rightarrow \mathcal{\dfiber}$\cite{barrCategoryTheoryComputing}.
\end{definition}

The object $\texttt{\dfiber}$ can be an object of any arbitrary category $\mathcal{C}$, which allows $\texttt{\dfiber}$ to encode most data fields. For example, a lists of strings is an instance of an object in \textbf{Set}, networks are an instance of \textbf{Graph}, and images are vector spaces which are a specific type of topological space \textbf{Top}.

%Monoid b/c arrows form a Monoid, F + Monoid that acts on F

%% F, M action on F, 
%%% F \->E->K

\begin{definition} For each open set object $U_{i}$ of category $\mathcal{\dbase}$, there is a set $\Gamma(U_i, \dtotal)$. The elements of this set are the sections $\{\tau^{j}:U_i \rightarrow \dfiber_{U_{i}}\}_{j\in I}$ of the fiber bundle \dtotal\ over open set $\U_{i}$.

  The sections $\Gamma(\dbase, \dtotal)$ are called the global sections of $\dtotal$. \cite{SheafMathematics2021,spanier1989algebraic}.
\end{definition}
Each section is how we represent a unique dataset, where the set of sections $\Gamma$ is the set of all datasets that have the same data fields in the \dfiber\ and continuity as encoded in\dbase\. For example, database tables that have the same schema and discrete index, as illustrated in Spivak \cite{spivakSIMPLICIALDATABASES}.

For example, these functions could be the binary operations, group actions, and measurement scales discussed in \autoref{sec:related-work:equivariance}. These functions could also be monoid actions, which are a way of applying partial order relations to data\cite{fongInvitationAppliedCategory2019}, such as to build multi-ranked indicators\cite{bruggemannRankingPrioritizationMultiindicator2011}

\begin{definition} The category $\mathcal{\dtotal}$ consists of 
  \begin{itemize}
    \item\textit{object} the space of continuous\footnote{or piecewise continuous}sections $\sheaf: U_{i} \rightarrow \Gamma(U_{i}, \dtotal) $ which is a sheaf\cite{spanier1989algebraic,SheafMathematics2021}. 
    \item\textit{morphisms} bundle maps E, E, fiber to itself and is the fiber map allowable, functions F to itself forms a monoid (one object category), F and it's allowable maps form a category, monoid could be monotonic functions, group actions, extending that to bundle E, interested in E to itself, which when restricted to fiber are category maps of fiber to itself, maps from sheaf to itself, that when restricted to fiber gives you fiber to itself, given bundle map E->E, then you can show that you have a sheaf map... Given E->E, and F->F, then there's a map from one sheaf to another. bundle map induces on level of sections induces on sheafs -> part of the definition of the presheaf
    \item \textit{morphisms} E->E such that restricted to fiber it's the allowed maps, and restricted to base it's the inclusion map, sheaf to itself so is an endofunctor (category to itself)
  \end{itemize}
\end{definition}

\begin{definition}We define the endofunctor \monoid on E as the structure preserving map which is the monoid. ...
\end{definition}

\note{equivariance is a special case of functions}
\note{actions are a special case of actions}
\begin{definition} category of bundle maps, which include category of fiber maps, what are allowable of fiber to itself, ex translation and scales only ..., objects are a fiber, maps are fiber to fiber defined as their own category, E to V fibers changes
\end{definition}

The objects of the data category $\mathcal{\dtotal}$ and graphic category $\mathcal{\gtotal}$ are the sheaves $\Gamma(\dbase, \dtotal)$ and $\Gamma(\gbase, \gtotal)$. An alternate notation is $\sheaf(\dtotal)$ and $\sheaf(\gtotal)$. The artist is a morphism of sheafs $\sheaf(\dtotal)$ to $\sheaf(\gtotal)$, which means it is by definition a natural transform\cite{SheafMathematics2021}. As mentioned in \autoref{sec:related-work:equivariance:category}
that the artist $\vartist$ transforms functions that act on data $f:\dtotal\rightarrow \dtotal$ to functions that act on graphics $\g: \gtotal\rightarrow \dtotal$ in a way in which the following diagram commutes 
\begin{equation}
  \begin{tikzcd}
    \mathcal{\dtotal} \arrow[r, "\vartist"] \arrow[d, "f"] & \mathcal{\gtotal} \arrow[d, "g"] \arrow[d, "\ca"] \\
    \mathcal{\dtotal} \arrow[r, "\vartist"]                & \mathcal{\gtotal}                                
  \end{tikzcd}
\end{equation}

meaning that, as described in \autoref{sec:related-work:equivariance}, the artist \vartist is constructed such that $\vartist(f(\dtotal)) = g(\vartist(\dtotal))$, meaning a transformation of the data has a consistent equivalent transformation of the graphic. In \autoref{sec:artist:construction}, we show a method of constructing the artist such that continuity and equivariance are preserved. We then use this formulation to guide the development of visualization library components in \autoref{sec:case-study}.

@@

% For this sake we don't do polymorphism, but we have a notion of equivalance classes of artists s.t... every parameter you want to fix is a fiber in E 
% given a nu, 
% maximal V has all the visual variables 
% nu_red()->red

@@

\subsection{Visualization Library Components}
\subsubsection{Graphic to Data: \vindex}
\begin{equation}
  \begin{tikzcd}
      \dtotal \arrow[d, "\pi"'] & \gtotal \arrow[d, "\pi"'] \\
      \dbase                   & \gbase \arrow[l, "\vindex"']
  \end{tikzcd}
  \label{eq:math:graphic:vindex}
\end{equation}

@@

\subsubsection{Visual Bundle \vtotal}
\begin{equation}
  \begin{tikzcd}[ampersand replacement=\&]
      \vfiber \arrow[r, hook] \& \vtotal \arrow[d, "\pi"'] \\
                        \& \dbase \arrow[u, "\vsection"', bend right]
  \end{tikzcd}
\end{equation}

\subsubsection{Data to Visual Encodings: \vchannel} %look at what K&S call these stages
Visual encoding functions \vindex\ are functors because they are expected to preserve structure on the data and visual side, as discussed in \autoref{sec:related-work:equivariance}.
